---
title: "All inclusive RNA code"
output: html_notebook
---

# Dependencies
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("DESeq2")   # run once if not installed
library(DESeq2)

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# install pcaExplorer
BiocManager::install("pcaExplorer")

# load it
library(pcaExplorer)

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("org.Hs.eg.db")
library(org.Hs.eg.db)

remove.packages("Hmisc")
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install(ask = FALSE, update = TRUE)
install.packages("Hmisc", dependencies = TRUE)
```

```{r}
library(tximport)
library(tidyverse)
library(Hmisc)
library(factoextra)
library(pheatmap)
library(ggrepel)
library(rhdf5)
library(matrixStats)
library(RColorBrewer)
library(GO.db)
library(topGO)
library(ComplexHeatmap)
library(ggplot2)
library(readr)
library(dplyr)
library(biomaRt)
library(cluster)
library(progeny)
#library(nichenetr)
library(singscore)
library(pheatmap)
library(tidyr)
library(viridisLite)
```

# Get transcript-to-gene annotation
```{r}
mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
  dataset = "hsapiens_gene_ensembl",
  host = 'https://www.ensembl.org')
t2g <- biomaRt::getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id",
    "external_gene_name"), mart = mart)
t2g <- dplyr::rename(t2g, target_id = ensembl_transcript_id,
  ens_gene = ensembl_gene_id, ext_gene = external_gene_name)
t2g <- unique(t2g[,c("target_id", "ext_gene")])
```


# Load the data
```{r}
sample_id <- dir("../data/rna_seq/")
kal_dirs <- file.path("../data/rna_seq", sample_id)
```
# Setup Annotation
```{r}
annotation <- data.frame(path = kal_dirs,
                         sample = sample_id,
                         cell_line = stringr::word(sample_id, 1, 1, sep="p")) #remove passage number
annotation$cell_line <- gsub("R", "-R", annotation$cell_line) #just adding dash before replicate info
```
# Change Adeno to HGSC
```{r}
subtype_dat <- read.csv("../data/human_sample_annotation_cleaned.csv", row.names=1)
subtype_dat$Subtype[subtype_dat$cell_line %in% c("OV3331", "OV90")] <- "High-grade serous"
write.csv(subtype_dat, "../output/annotation.csv")
```

```{r}
annotation$Subtype <- subtype_dat$Subtype[match(annotation$sample, subtype_dat$sample)]
```

## Import with tximport

```{r}
files <- file.path(annotation$path, "abundance.h5")
names(files) <- annotation$sample
```

```{r}
txi.kallisto <- tximport(files, type = "kallisto", txOut = FALSE, tx2gene = t2g,
                         ignoreTxVersion = T)
```

```{r}
tpm_mat <- txi.kallisto$abundance[2:nrow(txi.kallisto$abundance),] #the first row seems to be the sum of the transcripts missing from tx2gene
tpm_mat <- tpm_mat[rowSums(tpm_mat) != 0,] #remove genes with no detection in any sample'
#colnames(tpm_mat) <- annotation$cell_line
```

---------------
# Kallisto

```{r}
kallisto_output_dir <- "path/to/kallisto/output"
```

```{r}
files <- file.path(kallisto_output_dir, "abundance.tsv")
```

## Save point
```{r}
saveRDS(txi.kallisto, file="../output/txi.rds")
write.csv(tpm_mat, file="../output/tpm_matrix.csv", quote=F)
```
# Remove LGSC and set up sample annotation
```{r}

subtypes_to_exclude <- c("LGSOC_P1", "LGSOC_P10", "LGSOC_P11", 
                         "LGSOC_P2", "LGSOC_P3", "LGSOC_P4", 
                         "LGSOC_P5", "LGSOC_P7")
annotation <- annotation[!annotation$sample %in% subtypes_to_exclude, ]

# Filter TPM matrix columns (columns = cell line names)
tpm_mat <- tpm_mat[, colnames(tpm_mat) %in% annotation$cell_line]

# Update col_annotation to match
col_annotation <- data.frame(Subtype = annotation$Subtype)
rownames(col_annotation) <- annotation$cell_line

write.table(annotation, file = "../output/annotation.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
write.csv(annotation, file="../output/annotation.csv", quote=F)
```
# Reorder annotation table
```{r}
annotation <- read.csv("../output/annotation.csv", stringsAsFactors = FALSE)

subtype_order <- c("Clear cell", "Endometrioid", "High-grade serous", "MMMT", "Mucinous", "SCCOHT")

annotation$Subtype <- factor(annotation$Subtype, levels = subtype_order)

annotation <- annotation[order(annotation$Subtype), ]

write.csv(annotation, "../output/annotation.csv", row.names = FALSE)
write_tsv(annotation, "../output/annotation.tsv")

```
# High-level comparisons
```{r}
tpm_mat_log <- log1p(tpm_mat)
```

## Spearman correlation
```{r}
dat_cor <- cor(tpm_mat_log, method="spearman")

pdf("../figs2/spearman_cor_heatmap.pdf", width=6, height=5, units="in", res=300)


subtypes <- unique(annotation$Subtype)
subtype_colours <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
annotation_colours <- list(Subtype = subtype_colours)

pheatmap::pheatmap(dat_cor,
                   color = viridisLite::mako(100),
                   annotation_col = col_annotation,
                   annotation_colors = annotation_colours,
                   border_color = "black",
                   breaks = seq(0.825, 1, length.out=101),
                   show_colnames = FALSE,
                   clustering_method = "ward.D2",
                   fontsize = 9, 
                   filename = "../figs2/spearman_cor_heatmap.pdf", 
  width = 8, height = 6  
)

dev.off()
```

# QC Box plots
```{r}
qc_data <- read_tsv("../data/kallisto_qc.tsv")
annot <- read_tsv("../output/annotation.tsv")
annot <- read_csv("../output/annotation.csv")
annot_subset <- dplyr::select(annot, sample, Subtype)

qc_annot <- dplyr::left_join(qc_data, annot_subset, by = "sample")

subtypes <- unique(annot$Subtype)   
subtype_colours <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)

# Visualize number of processed reads
read_count_plot <- ggplot(qc_annot, aes(x = "n Processed Reads", y = n_processed)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90", width = 0.2) +
  geom_jitter(aes(color = Subtype), width = 0.1, size = 1.5, alpha = 0.9) +
  scale_color_manual(values = subtype_colours) +
  labs(x = "Number of Processed Reads", y = NULL , title = "Read count distribution across samples") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    plot.title = element_text(size = 10.5, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.position = "right"
  )
ggsave("../figs2/readcount_plot.pdf", read_count_plot, width = 2.79, height = 3, dpi = 300)


# Visualize number of unique counts
unique_count_plot <- ggplot(qc_annot, aes(x = "n_unique", y = n_unique)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90", width = 0.2) +
  geom_jitter(aes(color = Subtype), width = 0.1, size = 2, alpha = 0.9) +
  labs(x = "n_unique", y = NULL , title = "Unique values distribution across samples") +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    plot.title = element_text(size = 10.5, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.position = "right"
  )
ggsave("../figs2/uniquevalues_plot.pdf", unique_count_plot, width = 5, height = 4, dpi = 300)

# Visualize pseudoalignement rate
pseudoaligned_plot <- ggplot(qc_annot, aes(x = "Pseudoaligment rate (%)", y = p_pseudoaligned)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90", width = 0.2) +
  geom_jitter(aes(color = Subtype), width = 0.1, size = 1.5, alpha = 0.9) +
  scale_color_manual(values = subtype_colours) +
  labs(x = "Pseudoalignment rate (%)", y = NULL , title = "Pseudoaligned values distribution across samples") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    plot.title = element_text(size = 10.5, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.position = "right"
  )
ggsave("../figs2/pseudoalignedvalues_plot.pdf", pseudoaligned_plot, width = 2.78, height = 3.5, dpi = 300)
```

# tSNE
```{r}
ovcan_tsne <- Rtsne::Rtsne(pca_res$x[,1:5], perplexity=5) #keep 3*perp < nrow(x)
```

### Plotting
```{r}
plot_dat <- annotation
plot_dat$tSNE1 <- ovcan_tsne$Y[,1]
plot_dat$tSNE2 <- ovcan_tsne$Y[,2]
plot_dat$PC1 <- pca_res$x[,1]
plot_dat$PC2 <- pca_res$x[,2]
plot_dat$PC3 <- pca_res$x[,3]
plot_dat$PC4 <- pca_res$x[,4]
plot_dat$PC5 <- pca_res$x[,5]
plot_dat$PC6 <- pca_res$x[,6]
plot_dat$PC7 <- pca_res$x[,7]
plot_dat$PC8 <- pca_res$x[,8]
plot_dat$PC9 <- pca_res$x[,9]
plot_dat$PC10 <- pca_res$x[,10]

```
## Visualization
```{r}
subtypes <- unique(sub_anno$Subtype)
subtype_colours <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
anno_colours <- list(Subtype = subtype_colours)

tsne_plot <- ggplot(plot_dat, aes(x=tSNE1, y= tSNE2, label=cell_line)) + 
  geom_point(shape=21, size=2, color="black", aes(fill=Subtype)) +
  geom_text_repel(max.overlaps=100, size=1.75) +
  scale_fill_manual(values = subtype_colours) +
  theme_classic() +
  theme(axis.text=element_text(size=10, color="black"),
        axis.title=element_text(size=12))
ggsave("../figs2/tsne_plot_rna.pdf", plot = tsne_plot, width = 4.5, height = 2.75, dpi = 300)
```

# PCA prep
```{r}
subtypes <- unique(sub_anno$Subtype)
subtype_colours <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
anno_colours <- list(Subtype = subtype_colours)

tpm_mat_log_filtered <- tpm_mat_log[rowVars(as.matrix(tpm_mat_log)) > 0, ] #Remove genes with zero variance
pca_res <- prcomp(t(tpm_mat_log_filtered), scale. = TRUE)#Transpose for PCA (samples as rows, genes as columns)

# rownames(pca_res$x) <- colnames(tpm_mat_log_filtered) 



```
# PCA
```{r}
var_genes <- rownames(tpm_mat_log_filtered)[order(rowVars(tpm_mat_log_filtered),
                                        decreasing=T)[1:2000]]

tpm_mat_log_filtered <- tpm_mat_log_filtered[,annotation$cell_line]

pca_res <- prcomp(t(tpm_mat_log_filtered[var_genes,]), scale=F, center=T) 

# Base PCA plot
#fviz_pca_ind(pca_res, axes=c(1,2),
#              repel = TRUE,
#              habillage = annotation$Subtype,
#              max.overlaps=100)
```
## Visualization
```{r}
pca_plot <- ggplot(plot_dat, aes(x = PC1, y = PC2, label=cell_line)) + 
  geom_point(shape=21, size=2, color="black", aes(fill=Subtype)) +
  geom_text_repel(max.overlaps=100, size=1.75) +
  scale_fill_manual(values = subtype_colours) +
  theme_classic() +
  theme(axis.text=element_text(size=10, color="black"),
        axis.title=element_text(size=12)) + labs(x = "PC1 (20.1% Variance)", y = "PC2 (12.6% Variance)", fill = "Subtype")
ggsave("../figs2/pca_plot_rna.pdf", plot = pca_plot, width = 4.5, height = 2.75, dpi = 300)
```


#Sihouette Width PCA
```{r}
dists <- dist(plot_dat[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10")]) 

cluster_labels <- as.numeric(factor(plot_dat$Subtype))
sil <- silhouette(cluster_labels, dists) # basically subtype,distances
sil_df <- as.data.frame(sil)  # so i can use ggplot and dplyr
sil_df$Subtype <- plot_dat$Subtype # make subtype labels characters again

# average silhouette width per subtype
sil_summary <- sil_df %>%
  group_by(Subtype) %>%
  summarise(mean_sil = mean(sil_width)) %>%
  arrange(mean_sil)

# Visualization
sil_lollipop <- ggplot(sil_summary, aes(x = mean_sil, y = reorder(Subtype, mean_sil))) +
  geom_bar(stat = "identity", width = 0.01, color = "black") +
  geom_point(fill = "#4A154B", shape = 21, size = 4, color = "black") +
  geom_vline(xintercept = 0, linetype = 2) +
  ggtitle("Silhouette Width by Subtype (PCA)") +
  xlab("Mean Silhouette Width") + ylab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 26),
    plot.title = element_text(size = 26, face = "bold")
  )

ggsave("../figs2/silhouette_lollipop_1-10.pdf", plot = sil_lollipop, width = 7.5, height = 6)
```
# Euchlidean Distance Heatmap
```{r}

all_pcs <- (plot_dat[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10")]) 
rownames(all_pcs) <- plot_dat$cell_line  
dist_all <- dist(all_pcs, method = "euclidean")
invisible(dist_mat_all <- as.matrix(dist_all))

# Subtype annotation colouring
subtype_anno <- data.frame(Subtype = annotation$Subtype)
rownames(subtype_anno) <- annotation$cell_line
subtype_anno <- subtype_anno[rownames(dist_mat_all), , drop = FALSE]
subtypes <- unique(subtype_anno$Subtype)
subtype_colors <- setNames(
  RColorBrewer::brewer.pal(length(subtypes), "Set2"),
  subtypes)

# Visualization
euclidean_all_subtypes <- pheatmap::pheatmap(
  dist_mat_all,
  main = "Euclidean Distances (PC1 to PC10)",
  annotation_col = subtype_anno,
  annotation_colors = anno_colors,
  clustering_method = "complete",
  show_rownames = TRUE,
  show_colnames = TRUE,
  fontsize_row = 10,
  fontsize_col = 10,
  border_color = NA
)
ggsave("../figs2/euclidean_all_subtypes_annotation.pdf", plot = euclidean_all_subtypes, width = 10, height = 10)
```


# PCA on Clear cell lines only + Euclidean Distance Heatmap
```{r}

cc_samples <- plot_dat$cell_line[plot_dat$Subtype == "Clear cell"]

cc_expr <- tpm_mat_log_filtered[, cc_samples]
cc_expr_filtered <- cc_expr[rowVars(as.matrix(cc_expr)) > 0, ]

var_genes <- rownames(cc_expr_filtered)[order(rowVars(cc_expr_filtered),decreasing=T)[1:2000]]

cc_pca <- prcomp(t(cc_expr_filtered[var_genes,]), scale=T, center=T) 
cc_pcs <- cc_pca$x[, 1:4]
cc_dist <- dist(cc_pcs, method = "euclidean")
cc_mat <- as.matrix(cc_dist)


cc_heatmap <- pheatmap::pheatmap(cc_mat, main = "Clear Cell Euclidean Distances (PC1-4)")

ggsave("../figs2/clear_cell_euclidean_heatmap.pdf", plot = cc_heatmap)
```

# PCA on Mucinous lines only + Euclidean Distance Heatmap
```{r}

mc_samples <- plot_dat$cell_line[plot_dat$Subtype == "Mucinous"]

mc_expr <- tpm_mat_log_filtered[, mc_samples]
mc_expr_filtered <- mc_expr[rowVars(as.matrix(mc_expr)) > 0, ]
var_genes <- rownames(mc_expr_filtered)[order(rowVars(mc_expr_filtered),decreasing=T)[1:2000]]

mc_pca <- prcomp(t(mc_expr_filtered[var_genes,]), scale=T, center=T) 
mc_pcs <- mc_pca$x[, 1:2]
mc_dist <- dist(mc_pcs, method = "euclidean")
mc_mat <- as.matrix(mc_dist)

mc_heatmap <- pheatmap::pheatmap(mc_mat, main = "Mucinous Euclidean Distances (PC1-2)")

ggsave("../figs2/mucinous_euclidean_heatmap.pdf", plot = mc_heatmap)
```
# PCA on HGSC lines only + Euclidean Distance Heatmap
```{r}

hgs_samples <- plot_dat$cell_line[plot_dat$Subtype == "High-grade serous"]

hgs_expr <- tpm_mat_log_filtered[, hgs_samples]
hgs_expr_filtered <- hgs_expr[rowVars(as.matrix(hgs_expr)) > 0, ]
var_genes <- rownames(hgs_expr_filtered)[order(rowVars(hgs_expr_filtered),decreasing=T)[1:2000]]

hgs_pca <- prcomp(t(hgs_expr_filtered[var_genes,]), scale=T, center=T) 
hgs_pcs <- hgs_pca$x[, 1:8]
hgs_dist <- dist(hgs_pcs, method = "euclidean")
hgs_mat <- as.matrix(hgs_dist)

hgs_heatmap <- pheatmap::pheatmap(hgs_mat, main = "High-grade Euclidean Distances (PC1-8)")

ggsave("../figs2/hgs_euclidean_heatmap.pdf", plot = hgs_heatmap, height = 7)
```

# DESeq setup  
```{r}

# At this point, LGSC samples are already excluded from annotation and tpm_mat
# Ensure annotation is a clean factor
annotation$Subtype <- factor(annotation$Subtype)

# Filter txi.kallisto to include only the remaining (non-LGS) samples
txi_filtered <- list(
  counts = txi.kallisto$counts[, annotation$sample],
  abundance = txi.kallisto$abundance[, annotation$sample],
  length = txi.kallisto$length[, annotation$sample],
  countsFromAbundance = txi.kallisto$countsFromAbundance
)
```

```{r}
# countData = counts, colData = colData (your objects)

# 1) sanity check: same set of sample names?
stopifnot(setequal(colnames(counts), rownames(colData)))

# 2) reorder colData to match counts columns
colData <- colData[colnames(counts), , drop = FALSE]

# 3) now create the DESeq2 object
dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData   = colData,
  design    = ~ Subtype
)

dds <- DESeqDataSetFromTximport(txi_filtered, colData = annotations_filtered, design = ~ Subtype)
```
# Run DESeq and getting output files
```{r}
results_list <- list()
for (subtype in unique(annotation$Subtype)) {
    # above is the subtype loop for each unique subtype
  #makes a true vs false loop. if true itll be the subtype if false itll be the group
    annotation$Group <- factor(ifelse(annotation$Subtype == subtype, subtype, "Other"))
    # run the deseq for 
    dds <- DESeqDataSetFromTximport(txi_filtered, colData = annotation, design = ~ Group)
    dds <- DESeq(dds)
    # get results subtype vs all others
    res <- results(dds, contrast = c("Group", subtype, "Other"))
    results_list[[subtype]] <- res
    output_filename <- paste0("differential_expression_", subtype, "_vs_Other.csv")
    write.csv(as.data.frame(res), file = output_filename)}
```

# Signature for HGSC with pajd/log2fc
```{r}
hgs <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_HGS_vs_Other.csv", row.names = 1)
hgs_signature_df <- subset(hgs, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
hgs_signature_df$Gene <- rownames(hgs_signature_df)
hgs_signature_df <- hgs_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(hgs_signature_df,
          "../output/HGS_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)
```
## Go terms HGSC
```{r}
gene_list <- na.omit(gene_list)
go_up_hgs <- pcaExplorer::topGOtable(hgs_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_hgs
```
### Visualize 
```{r}
up_labels_hgs <- c(
  "cellular response to hypoxia",
  "telomere organization",
  "nucleosome assembly",
  "double-strand break repair",
  "proton motive force-driven mitochondrial ATP synthesis"
)

df_up_hgs <- go_up_hgs %>%
  filter(Term %in% up_labels_hgs)
df_up_hgs$Term <- Hmisc::capitalize(df_up_hgs$Term)
df_up_hgs$Term <- factor(df_up_hgs$Term, levels = rev(df_up_hgs$Term))

# Visualize top GO terms
goterms_hgs <- ggplot(df_up_hgs, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("HGS GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
  theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))
ggsave("../figs2/HGS_goterms.pdf", plot = goterms_hgs, width = 10, height = 2.5)

# all go terms in a csv file
#write.csv(go_up_hgs, "../output/HGS_goterms_all.csv", row.names = FALSE)

```

#### Volcano Plot
```{r}
# Categorize genes based on fold change and p-value
hgs$diffexpressed <- "NO"
hgs$diffexpressed[hgs$log2FoldChange > 0.6 & hgs$pvalue < 0.05] <- "UP"
hgs$diffexpressed[hgs$log2FoldChange < -0.6 & hgs$pvalue < 0.05] <- "DOWN"

# Label only the top 30 most significant UP-regulated genes (based on padj)
up_genes <- hgs %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)

top_up <- head(up_genes, 20)
hgs$delabel <- ifelse(hgs$X %in% top_up, hgs$X, NA) # Add a new column 'delabel' that keeps the gene name only for the top 20 upregulated genes; others are NA (unlabeled)

# Create the volcano plot
hgs_volcano <- ggplot(data = hgs, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("HGS Significant Genes") +
  scale_color_manual(values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
                     labels = c("Downregulated", "Not significant", "Upregulated")) +
  coord_cartesian(ylim = c(0, 20), xlim = c(-10, 10)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
  theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  )

# Save to PDF
pdf(file = "../figs2/hgs_volcano_rna.pdf", width = 8, height = 8)
hgs_volcano
dev.off() 
```


# Signature for Mucinous with pajd/log2fc
```{r}
mc <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_Mucinous_vs_Other.csv") #, row.names = 1)
mc_signature_df <- subset(mc, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
mc_signature_df$Gene <- rownames(mc_signature_df)
mc_signature_df <- mc_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(mc_signature_df,
          "../output/Mucinous_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)



```
##Go terms Mucinous
```{r}
gene_list <- na.omit(gene_list)
go_up_mc <- pcaExplorer::topGOtable(mc_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_mc
```
###Visualize 
```{r}
up_labels_mc <- c(
  "complement activation, classical pathway",
  "oligosaccharide biosynthetic process",
  "cellular lipid metabolic process",
  "regulation of stress-activated MAPK cascade",
  "extracellular matrix organization"
)

df_up_mc <- go_up_mc %>%
  filter(Term %in% up_labels_mc)
df_up_mc$Term <- Hmisc::capitalize(df_up_mc$Term)
df_up_mc$Term <- factor(df_up_mc$Term, levels = rev(df_up_mc$Term))

# Visualize top GO terms
goterms_mc <- ggplot(df_up_mc, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("Mucinous GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
    theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))

ggsave("../figs2/MC_goterms.pdf", plot = goterms_mc, width = 10, height = 2.5)
#write.csv(go_up_mc, "../output/MC_goterms_all.csv", row.names = FALSE)
```
#### Volcano Plot
```{r}
mc$diffexpressed <- "NO"
mc$diffexpressed[mc$log2FoldChange > 0.6 & mc$pvalue < 0.05] <- "UP"
mc$diffexpressed[mc$log2FoldChange < -0.6 & mc$pvalue < 0.05] <- "DOWN"

mc_up <- mc %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)
mc_top_up <- head(mc_up, 20)
mc$delabel <- ifelse(mc$X %in% mc_top_up, mc$X, NA)

mc_volcano <- ggplot(mc, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("Mucinous Significant Genes") +
  scale_color_manual(
  values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
  labels = c("DOWN" = "Downregulated", "NO" = "Not Significant", "UP" = "Upregulated")
) + coord_cartesian(ylim = c(0, 13), xlim = c(-10, 10)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
   theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  
  )
# Save
pdf(file = "../figs2/RNA_volcano_plots/mc_volcano.pdf", width = 8, height = 8)
mc_volcano
dev.off()
```
# Signature for Clear Cell with pajd/log2fc
```{r}
cc <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_Clear cell_vs_Other.csv") #, row.names = 1)
cc_signature_df <- subset(cc, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
cc_signature_df$Gene <- rownames(cc_signature_df)
cc_signature_df <- cc_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(cc_signature_df,
          "../output/ClearCell_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)

```
##Go terms Clear Cell
```{r}
gene_list <- na.omit(gene_list)
go_up_cc <- pcaExplorer::topGOtable(cc_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_cc
```
###Visualize 
```{r}
up_labels_cc <- c(
  "response to oxidative stress",
  "basement membrane organization",
  "positive regulation of receptor signaling pathway via JAK-STAT",
  "negative regulation of stress-activated MAPK cascade",
  "glutathione catabolic process"
)

# Filter and Visualize top GO terms
df_up_cc <- go_up_cc %>%
  filter(Term %in% up_labels_cc)
df_up_cc$Term <- Hmisc::capitalize(df_up_cc$Term)
df_up_cc$Term <- factor(df_up_cc$Term, levels = rev(df_up_cc$Term))
goterms_cc <- ggplot(df_up_cc, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("Clear Cell GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
   theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))
ggsave("../figs2/CC_goterms.pdf", plot = goterms_cc, width = 12, height = 2.5)
#write.csv(go_up_cc, "../output/CC_goterms_all.csv", row.names = FALSE)
```
#### Volcano Plot
```{r}
# Categorize expression
cc$diffexpressed <- "NO"
cc$diffexpressed[cc$log2FoldChange > 0.6 & cc$pvalue < 0.05] <- "UP"
cc$diffexpressed[cc$log2FoldChange < -0.6 & cc$pvalue < 0.05] <- "DOWN"

# Label top 30 UP-regulated genes only
cc_up <- cc %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)
cc_top_up <- head(cc_up, 20)
cc$delabel <- ifelse(cc$X %in% cc_top_up, cc$X, NA)

# Volcano plot
cc_volcano <- ggplot(cc, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("CC Significant Genes") +
   scale_color_manual(values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
                     labels = c("Downregulated", "Not significant", "Upregulated")) +
  coord_cartesian(ylim = c(0, 20), xlim = c(-10, 10)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
   theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  )


# Save
pdf(file = "../figs2/RNA_volcano_plots/cc_volcano.pdf", width = 8, height = 8)
cc_volcano
dev.off()

```
# Signature for Endometrioid with pajd/log2fc
```{r}
ec <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_Endometrioid_vs_Other.csv", row.names = 1)
ec_signature_df <- subset(ec, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
ec_signature_df$Gene <- rownames(ec_signature_df)
ec_signature_df <- ec_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(ec_signature_df,
          "../output/ec_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)

```
##Go terms Endometrioid
```{r}
gene_list <- na.omit(gene_list)
go_up_ec <- pcaExplorer::topGOtable(ec_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_ec
```
###Visualize 
```{r}
up_labels_ec <- c(
  "positive regulation of canonical Wnt signaling pathway",
  "regulation of epithelial cell proliferation",
  "collagen fibril organization",
  "extracellular matrix assembly",
  "negative regulation of transcription by RNA polymerase II"
)

# Filter and prep
df_up_ec <- go_up_ec %>%
  filter(Term %in% up_labels_ec)
df_up_ec$Term <- Hmisc::capitalize(df_up_ec$Term)
df_up_ec$Term <- factor(df_up_ec$Term, levels = rev(df_up_ec$Term))
# Visualize top GO terms
goterms_ec <- ggplot(df_up_ec, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("Endometrioid GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
   theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))
ggsave("../figs2/ec_goterms.pdf", plot = goterms_ec, width = 12, height = 2.5)

# save just csv file with everything
#write.csv(go_up_ec, "../output/EC_goterms_all.csv", row.names = FALSE)
```
#### Volcano Plot
```{r}
ec$diffexpressed <- "NO"
ec$diffexpressed[ec$log2FoldChange > 0.6 & ec$pvalue < 0.05] <- "UP"
ec$diffexpressed[ec$log2FoldChange < -0.6 & ec$pvalue < 0.05] <- "DOWN"

ec_up <- ec %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)
ec_top_up <- head(ec_up, 20)
ec$delabel <- ifelse(ec$X %in% ec_top_up, ec$X, NA)

ec_volcano <- ggplot(ec, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("Endometrioid Significant Genes") +
  scale_color_manual(
  values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
  labels = c("DOWN" = "Downregulated", "NO" = "Not Significant", "UP" = "Upregulated")
) +
  coord_cartesian(ylim = c(0, 20), xlim = c(-15, 15)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
   theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  )

pdf(file = "../figs2/RNA_volcano_plots/ec_volcano.pdf", width = 8, height = 8)
ec_volcano
dev.off()

```

# Signature for SCCOHT with pajd/log2fc
```{r}
sccoht <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_SCCOHT_vs_Other.csv", row.names = 1)
sccoht_signature_df <- subset(sccoht, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
sccoht_signature_df$Gene <- rownames(sccoht_signature_df)
sccoht_signature_df <- sccoht_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(sccoht_signature_df,
          "../output/sccoht_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)

```
##Go terms SCCOHT
```{r}
gene_list <- na.omit(gene_list)
go_up_sccoht <- pcaExplorer::topGOtable(sccoht_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_sccoht

```
###Visualize 
```{r}
up_labels_sccoht <- go_up_sccoht$Term[1:5]
df_up_sccoht <- go_up_sccoht %>%
  filter(Term %in% up_labels_sccoht)
df_up_sccoht$Term <- Hmisc::capitalize(df_up_sccoht$Term)
df_up_sccoht$Term <- factor(df_up_sccoht$Term, levels=rev(df_up_sccoht$Term))
bluered <- rev(RColorBrewer::brewer.pal(9, "RdBu"))

# Visualize top GO terms
goterms_sccoht <- ggplot(df_up_sccoht, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("SCCOHT GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
 theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))
ggsave("../figs2/sccoht_goterms.pdf", plot = goterms_sccoht, width = 12, height = 2.5)

# save just csv file with everything
#write.csv(go_up_sccoht, "../output/Go_terms/SCCOHT_goterms_all.csv", row.names = FALSE)
```
#### Volcano Plot
```{r}
# Categorize expression status
sccoht$diffexpressed <- "NO"
sccoht$diffexpressed[sccoht$log2FoldChange > 0.6 & sccoht$pvalue < 0.05] <- "UP"
sccoht$diffexpressed[sccoht$log2FoldChange < -0.6 & sccoht$pvalue < 0.05] <- "DOWN"

# Top 30 most significant UP genes by adjusted p-value
sccoht_up <- sccoht %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)

sccoht_top_up <- head(sccoht_up, 20)
sccoht$delabel <- ifelse(sccoht$X %in% sccoht_top_up, sccoht$X, NA)

# Plot
sccoht_volcano <- ggplot(sccoht, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("SCCOHT Significant Genes") +
  scale_color_manual(
    values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
    labels = c("DOWN" = "Downregulated", "NO" = "Not Significant", "UP" = "Upregulated")
  ) +
  coord_cartesian(ylim = c(0, 20), xlim = c(-12, 12)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
   theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  )

# Save
pdf(file = "../figs2/RNA_volcano_plots/sccoht_volcano.pdf", width = 8, height = 8)
sccoht_volcano
dev.off()

```


# Signature for MMMT with pajd/log2fc
```{r}
mmmt <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/differential_expression_MMMT_vs_Other.csv", row.names = 1)
mmmt_signature_df <- subset(mmmt, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
mmmt_signature_df$Gene <- rownames(mmmt_signature_df)
mmmt_signature_df <- mmmt_signature_df[, c("Gene", "padj", "log2FoldChange")]
write.csv(mmmt_signature_df,
          "../output/MMMT_signature_detailed.csv",
          row.names = FALSE,
          quote = FALSE)

```
##Go terms MMMT
```{r}
gene_list <- na.omit(gene_list)
go_up_mmmt <- pcaExplorer::topGOtable(mmmt_signature_df %>%  
                                   filter(padj < 0.05 & log2FoldChange > 1) %>%
                                   pull(Gene),
                                rownames(dds),
                                ontology="BP",
                                mapping="org.Hs.eg.db")

```
```{r}
go_up_mmmt
```
###Visualize 
```{r}
up_labels_mmmt <- go_up_mmmt$Term[1:5]
df_up_mmmt <- go_up_mmmt %>%
  filter(Term %in% up_labels_mmmt)
df_up_mmmt$Term <- Hmisc::capitalize(df_up_mmmt$Term)
df_up_mmmt$Term <- factor(df_up_mmmt$Term, levels=rev(df_up_mmmt$Term))
bluered <- rev(RColorBrewer::brewer.pal(9, "RdBu"))
# Visualize top GO terms
goterms_mmmt <- ggplot(df_up_mmmt, aes(x=-log10(p.value_elim), y=Term)) +
  geom_bar(stat="identity", width = 0.01, color="black") +
  geom_point(fill="#4A154B", shape=21, size=4, color="black") +
  geom_vline(xintercept = -log10(0.05), linetype=2) +
  ggtitle("MMMT GO Terms") +
  xlab("-log10(padj)") + ylab("") +
  theme_classic() +
  theme(legend.position="none",
        axis.text=element_text(size=20, color="black"),
        axis.title=element_text(size=20),
        plot.title = element_text(size = 26, face = "bold"))
ggsave("../figs2/MMMT_goterms_2026.pdf", plot = goterms_mmmt, width = 12, height = 2.5)

# save just csv file with everything
write.csv(go_up_mmmt, "../output/MMMT_goterms_all.csv", row.names = FALSE)
```
#### Volcano plot
```{r}
mmmt$diffexpressed <- "NO"
mmmt$diffexpressed[mmmt$log2FoldChange > 0.6 & mmmt$pvalue < 0.05] <- "UP"
mmmt$diffexpressed[mmmt$log2FoldChange < -0.6 & mmmt$pvalue < 0.05] <- "DOWN"

mmmt_up <- mmmt %>%
  filter(diffexpressed == "UP" & !is.na(padj)) %>%
  arrange(padj) %>%
  pull(X)
mmmt_top_up <- head(mmmt_up, 20)
mmmt$delabel <- ifelse(mmmt$X %in% mmmt_top_up, mmmt$X, NA)

mmmt_volcano <- ggplot(mmmt, aes(x = log2FoldChange, y = -log10(pvalue), col = diffexpressed, label = delabel)) +
  geom_point() +
  ggtitle("MMMT Significant Genes") +
  scale_color_manual(
    values = c("DOWN" = "#00AFBB", "NO" = "grey", "UP" = "red"),
    labels = c("DOWN" = "Downregulated", "NO" = "Not Significant", "UP" = "Upregulated")
  ) +
  coord_cartesian(ylim = c(0, 20), xlim = c(-15, 15)) +
  geom_text_repel(max.overlaps = Inf, na.rm = TRUE, size = 5) + theme_classic() +
  theme(
    axis.title = element_text(size = 29),
    axis.text = element_text(size = 29),
    plot.title = element_text(size = 29, face = "bold"),
    legend.title = element_text(size = 23),
    legend.text = element_text(size = 20)
  )

pdf(file = "../figs2/RNA_volcano_plots/mmmt_volcano.pdf", width = 8, height = 8)
mmmt_volcano
dev.off()
```


# Signature Heatmap top 10

```{r}
tpm_mat_log <- tpm_mat_log[rowVars(as.matrix(tpm_mat_log)) != 0, ]
tpm_mat_log <- t(scale(t(tpm_mat_log))) 

# load and identify all complete gene signatures for each subtype
hgs_sig <- read.csv("../output/HGS_signature_detailed.csv")
hgs_sig$Subtype <- "High-grade serous"

mc_sig <- read.csv("../output/Mucinous_signature_detailed.csv")
mc_sig$Subtype <- "Mucinous"

ec_sig <- read.csv("../output/ec_signature_detailed.csv")
ec_sig$Subtype <- "Endometrioid"

cc_sig <- read.csv("../output/ClearCell_signature_detailed.csv")
cc_sig$Subtype <- "Clear cell"

mmmt_sig <- read.csv("../output/MMMT_signature_detailed.csv")
mmmt_sig$Subtype <- "MMMT"

sccoht_sig <- read.csv("../output/SCCOHT_signature_detailed.csv")
sccoht_sig$Subtype <- "SCCOHT"

# join together into one big table with top 10 genes for each subtype
df_sig_hm <- bind_rows(hgs_sig, mc_sig, ec_sig, cc_sig, mmmt_sig, sccoht_sig) %>%
  group_by(Subtype) %>%
  slice_min(order_by = padj, n = 10, with_ties = FALSE)

# Filter genes present in matrix and order by subtype
df_sig_hm <- df_sig_hm %>%
  filter(Gene %in% rownames(tpm_mat_log)) %>%
  arrange(Subtype, padj)

ordered_genes <- df_sig_hm$Gene

tpm_subset <- tpm_mat_log[ordered_genes, ]
tpm_subset[tpm_subset > 2] <- 2
tpm_subset[tpm_subset < -2] <- -2

sample_anno <- read.csv("../output/annotation.csv") %>%
  filter(cell_line %in% colnames(tpm_subset)) %>%
  arrange(Subtype)

# columns in matrix to match subtype order
tpm_subset <- tpm_subset[, sample_anno$cell_line]

# Build row subtype vector
subtype_vector_row <- df_sig_hm$Subtype
names(subtype_vector_row) <- df_sig_hm$Gene  

# subtype color palette
unique_subtypes <- sort(unique(subtype_vector_row))
subtype_colors <- setNames(
  RColorBrewer::brewer.pal(length(unique_subtypes), "Set2"),
  unique_subtypes)

# row annotation
row_anno <- rowAnnotation(
  Subtype = subtype_vector_row[rownames(tpm_subset)],
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE)

# bottom annotation
bottom_anno <- HeatmapAnnotation(
  Subtype = sample_anno$Subtype,
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE
)

sigs_heatmap <- Heatmap(
  tpm_subset,
  name = "Expression\nZ-score",
  show_row_names = F,
  show_column_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 6),
  heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
  border = TRUE,
  left_annotation = row_anno,
  bottom_annotation = bottom_anno)

 pdf(file = "../figs2/sigs_heatmap_rna_top10.pdf", width = 8, height = 8)
draw(sigs_heatmap)
dev.off()

```

# Signature Heatmap top 100 
(same as above, just top 100 now)
```{r}
tpm_mat_log <- tpm_mat_log[rowVars(as.matrix(tpm_mat_log)) != 0, ]
tpm_mat_log <- t(scale(t(tpm_mat_log)))  # Z-score per gene

# Load signature files and assign subtypes
hgs_sig <- read.csv("../output/HGS_signature_detailed.csv")
hgs_sig$Subtype <- "High-grade serous"

mc_sig <- read.csv("../output/Mucinous_signature_detailed.csv")
mc_sig$Subtype <- "Mucinous"

ec_sig <- read.csv("../output/ec_signature_detailed.csv")
ec_sig$Subtype <- "Endometrioid"

cc_sig <- read.csv("../output/ClearCell_signature_detailed.csv")
cc_sig$Subtype <- "Clear cell"

mmmt_sig <- read.csv("../output/MMMT_signature_detailed.csv")
mmmt_sig$Subtype <- "MMMT"

sccoht_sig <- read.csv("../output/SCCOHT_signature_detailed.csv")
sccoht_sig$Subtype <- "SCCOHT"

# Combine and select top 100 DEGs per subtype
df_sig_hm <- bind_rows(hgs_sig, mc_sig, ec_sig, cc_sig, mmmt_sig, sccoht_sig) %>%
  group_by(Subtype) %>%
  slice_min(order_by = padj, n = 100, with_ties = FALSE)

# Filter genes present in matrix and order by subtype
df_sig_hm <- df_sig_hm %>%
  filter(Gene %in% rownames(tpm_mat_log)) %>%
  arrange(Subtype, padj)

# Get ordered gene list
ordered_genes <- df_sig_hm$Gene

# Subset and scale TPM matrix
tpm_subset <- tpm_mat_log[ordered_genes, ]
tpm_subset[tpm_subset > 2] <- 2
tpm_subset[tpm_subset < -2] <- -2

sample_anno <- read.csv("../output/annotation.csv") %>%
  filter(cell_line %in% colnames(tpm_subset)) %>%
  arrange(Subtype)

# Reorder columns in matrix to match subtype order
tpm_subset <- tpm_subset[, sample_anno$cell_line]

# Build row subtype vector
subtype_vector_row <- df_sig_hm$Subtype
names(subtype_vector_row) <- df_sig_hm$Gene  # ensure names match rownames(tpm_subset)

# Create subtype color palette
unique_subtypes <- sort(unique(subtype_vector_row))
subtype_colors <- setNames(
  RColorBrewer::brewer.pal(length(unique_subtypes), "Set2"),
  unique_subtypes
)

# Build row annotation
row_anno <- rowAnnotation(
  Subtype = subtype_vector_row[rownames(tpm_subset)],
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE
)

bottom_anno <- HeatmapAnnotation(
  Subtype = sample_anno$Subtype,
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE
)
# Draw heatmap with row annotation
sigs_heatmap <- Heatmap(
  tpm_subset,
  name = "Expression\nZ-score",
  show_row_names = F,
  show_column_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 10),
  row_names_gp = gpar(fontsize = 6),
  heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
  border = TRUE,
  left_annotation = row_anno,
  bottom_annotation = bottom_anno
)

pdf(file = "../figs2/RNAsigs_heatmap_top100.pdf", width = 8, height = 8)
draw(sigs_heatmap)
dev.off()
```
# Gene expression plots
```{r}
# Function to plot specific gene expression across subtypes
plot_gene_expression <- function(gene) {
  if (!(gene %in% rownames(tpm_mat_log))) {
    stop("The specified gene is not present in the expression matrix.")
  }

  gene_expression <- tpm_mat_log[gene, ] # find the row of the gene I called for and take all the columns for that gene

  plot_data <- data.frame(
    Subtype = annotation$Subtype,
    Expression = gene_expression
  )

  p <- ggplot(plot_data, aes(x = Subtype, y = Expression)) +
    geom_jitter(width = 0.2, color = "#0072B2", size = 1.5, alpha = 0.7) +
    theme_minimal() +
    labs(
      title = paste("Gene Expression for", gene),
      x = "Subtype",
      y = "Log(TPM + 1)"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme_classic()

  return(p)
}

# Plot gene
gene_to_plot <- "A1BG"
gene_plot <- plot_gene_expression(gene_to_plot)

# Display and save
print(gene_plot)
ggsave(filename = paste0("../figs2/Gene Expression/", gene_to_plot, "_rnaexp.pdf"),
       plot = gene_plot, width = 8, height = 4.5)

```

# PROGENy Heatmap all subtypes
## Gene sets
```{r}
#Hallmark EMT
hallmarks <- fgsea::gmtPathways("../data/h.all.v7.4.symbols.gmt.txt")
emt_hallmark <- hallmarks[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION"]]
```
##Some more hallmarks
```{r}
angiogenesis <- hallmarks[["HALLMARK_ANGIOGENESIS"]]

cholesterol <- hallmarks[["HALLMARK_CHOLESTEROL_HOMEOSTASIS"]]

glycolysis <- hallmarks[["HALLMARK_GLYCOLYSIS"]]

mtorc <- hallmarks[["HALLMARK_MTORC1_SIGNALING"]]

myc <- hallmarks[["HALLMARK_MYC_TARGETS_V2"]]

oxphos <- hallmarks[["HALLMARK_OXIDATIVE_PHOSPHORYLATION"]]

kras <- hallmarks[["HALLMARK_KRAS_SIGNALING_UP"]]

ros <- hallmarks[["HALLMARK_REACTIVE_OXYGEN_SPECIES_PATHWAY"]]

e2f <- hallmarks[["HALLMARK_E2F_TARGETS"]]

hallmarks <- list("EMT" = emt_hallmark,
                  "Angiogenesis" = angiogenesis,
                  "Cholesterol Metabolism" = cholesterol,
                  "Glycolysis" = glycolysis,
                  "MTORC1 Signalling" = mtorc,
                  "Myc Signalling" = myc,
                  "Oxidative Phosphorylation" = oxphos,
                  "KRAS Signalling" = kras,
                  "ROS Signalling" = ros,
                  "E2F Targets" = e2f)
pdl1 <- "CD274"
mhc1a <- c("HLA-A", "HLA-B", "HLA-C")
```
### YAP
```{r}
# YAP Targets from Wang et al., Cell Reports, 2018 ("Comprehensive molecular characterization of the hippo signaling pathway in cancer")
yap_targets <- c("AMOTL2", "ANKRD1", "IGFBP3", "F3", "FJX1",
                 "NUAK2", "LATS2", "CRIM1", "GADD45A", "TGFB2", "PTPN14", "NT5E",
                 "FOXF2", "AXL", "DOCK5", "ASAP1", "RBMS3", "MYOF", "ARHGEF17",
                 "CCDC80")
#yap_targets <- convert_human_to_mouse_symbols(yap_targets) %>% na.omit() %>% as.character()
yap_targets <- c(yap_targets, "CCN1", "CCN2") #Ctgf and Cyr61 oddly not in the nichenet converter???

#Chemoresistance in ovarian cancer
# From Sun et al, Cell Death and Disease, 2019
chemoresistance <- c("EDIL3", "GNG12", "MBOAT2", "MTMR6", "NBR1", "NEK7", 
                     "NET1", "PPP3CA", "RAD17", "WDR41")

chemosensitive <- c("FZD4", "MUTYH", "PCK2", "PEX10", "SRPK1", "UCP2")

#HRD
hrd <- read.table("../data/ovarian_cancer_HRD_peng_NatComm",
                  header=T)
hrd <- colnames(hrd)
```


```{r}
# Han et al metagenes
genesets <- read.csv("../data/Han_NatComm_2020_Signalling_Metagenes_Cleaned.csv")

han <- list()
for(i in 1:ncol(genesets)){
  genes <- genesets[,i]
  genes <- na.omit(convert_mouse_to_human_symbols(genes))
  han[[i]] <- genes[genes != ""]
}
names(han) <- colnames(genesets)

#Subset to only keep targets
han_targets <- han[grep("Targets", names(han), value=T)]
han_ligands <- han[grep("Ligands", names(han), value=T)]
#han_ligands <- convert_mouse_to_human_symbols(han_targets)
```

##Combine
```{r}

genesets <- list("PD-L1" = pdl1, 
              "MHC-Ia" = mhc1a, 
              "YAP Targets" = yap_targets,
              "Chemoresistance" = chemoresistance,
              "Chemosensitivity" = chemosensitive,
              "HRD" = hrd)
genesets <- c(genesets, hallmarks, han_targets)
names(genesets) <- gsub("_", " ", names(genesets))
```
## Gene Set Scoring

```{r}
tpm_ranked <- rankGenes(tpm_mat)

score_res <- list()

for(i in 1:length(genesets)){
  print(paste0("Scoring: ", names(genesets)[i]))
  
  res <- simpleScore(tpm_ranked, upSet=unique(genesets[[i]]))
  score_res[[i]] <- data.frame(SampleID = annotation$cell_line,
                               GeneSet = names(genesets)[i],
                               Score = res$TotalScore)
}

score_res <- do.call("rbind", score_res)
```

```{r}
score_res$Subtype <- annotation$Subtype[match(score_res$SampleID, annotation$cell_line)]
```
```{r}

score_summary <- score_res %>%
  group_by(GeneSet) %>%
  mutate(Score_Z = (Score - mean(Score)) / sd(Score))

```
##progeny 
```{r}
mat <- log1p(as.matrix(tpm_mat))

pathways <- progeny(mat, organism = "Human", top = 500)

pathway_summary_new <- pathways %>%
  as.data.frame() %>%
  rownames_to_column(var = "SampleID") %>%
  mutate(subtype = annotation$Subtype) %>%
  pivot_longer(-c(SampleID, subtype), names_to = "pathways", values_to = "Score") %>%
  group_by(pathways) %>%
  mutate(Activity_Z = (Score - mean(Score)) / sd(Score))
  
```

```{r}
score_matrix <- score_summary %>%
  dplyr::select(GeneSet, Score_Z, SampleID) %>%
  pivot_wider(names_from = GeneSet, values_from = Score_Z) %>%
  column_to_rownames(var = "SampleID") %>%
  as.matrix()

pathway_matrix <- pathway_summary_new %>%
  dplyr::select(pathways, Activity_Z, SampleID) %>%
  pivot_wider(names_from = pathways, values_from = Activity_Z) %>%
  column_to_rownames(var = "SampleID") %>%
  as.matrix()
```

# Visualization
```{r}

line_order <- annotation %>%
  arrange(Subtype) %>%
  pull(cell_line)
annotation_ordered <- annotation[annotation$cell_line %in% rownames(score_matrix), ]

# order the samples based on subtype
ordered_samples <- annotation_ordered$cell_line[order(annotation_ordered$Subtype)]

score_matrix <- score_matrix[ordered_samples, ]

subtype_anno <- rowAnnotation(
  Subtype = annotation_ordered$Subtype[match(rownames(score_matrix), annotation_ordered$cell_line)],

  col = list(Subtype = c(
    "Clear cell" = "#5EB5A6",
    "Endometrioid" = "#FF6F61",
    "High-grade serous" = "#8A9EC2",
    "MMMT" = "#CC79A7",
    "Mucinous" = "#A3D55F",
    "SCCOHT" = "#F0E442"
  )),
  annotation_name_gp = gpar(fontsize = 10),
  annotation_legend_param = list(title_gp = gpar(fontsize = 10)),
  gap = unit(2, "mm") 
)


h1 <- Heatmap(pathway_matrix,
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = T,
              cluster_rows = F,
              show_column_dend = T,
              clustering_method_rows = "ward.D2",
              show_row_names = T,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(pathway_matrix) * unit(5, "mm"), 
              height = nrow(pathway_matrix) * unit(5, "mm"))

h2 <- Heatmap(score_matrix[, meta_pathway],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = T,
              cluster_rows = F,
              show_column_dend = T,
              clustering_method_rows = "ward.D2",
              show_row_names = T,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(score_matrix[, meta_pathway]) * unit(5, "mm"), 
              height = nrow(score_matrix[, meta_pathway]) * unit(5, "mm"))

h3 <- Heatmap(score_matrix[, functions],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = T,
              cluster_rows = F,
              show_column_dend = T,
              clustering_method_rows = "ward.D2",
              show_row_names = T,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(score_matrix[, functions]) * unit(5, "mm"), 
              height = nrow(score_matrix[, functions]) * unit(5, "mm"))

h4 <- Heatmap(score_matrix[, metabolism],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = T,
              cluster_rows = F,
              show_column_dend = T,
              clustering_method_rows = "ward.D2",
              show_row_names = T,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(score_matrix[, metabolism]) * unit(5, "mm"), 
              height = nrow(score_matrix[, metabolism]) * unit(5, "mm"))
ht_list <- subtype_anno + h1 + h2 + h3 + h4 
pdf(file = "../figs2/heatmap_allsubtypes.pdf", width = 15, height = 15)
draw(ht_list)
dev.off()
```


# PROGENy Heatmap HGS
## Scaling after subsetting by subtype  HGS
```{r}
hgsc_lines <- annotation$cell_line[annotation$Subtype == "High-grade serous"]

#scaling
#Subset score_res to HGS
score_res_hgs <- score_res %>%
  filter(SampleID %in% hgsc_lines)

#scale matrix
filtered_score_summary_hgs <- score_res_hgs %>%
  group_by(GeneSet) %>%
  mutate(Score_Z = scale(Score))%>%
  ungroup()


#score matrix
filtered_score_matrix_hgs <- filtered_score_summary_hgs %>%
  dplyr::select(GeneSet, Score_Z, SampleID) %>%
  pivot_wider(names_from = GeneSet, values_from = Score_Z) %>%
  column_to_rownames("SampleID") %>%
  as.matrix()


#scale pathway
#fix pathway rownames
rownames(pathways) <- gsub("\\.", "-", rownames(pathways))
hgsc_lines <- hgsc_lines[hgsc_lines %in% rownames(pathways)]
pathways_hgs <- pathways[hgsc_lines, ]
#score pathways
filtered_pathway_summary_hgs <- pathways_hgs %>%
  as.data.frame() %>%
  rownames_to_column("SampleID") %>%
  pivot_longer(cols = -SampleID, names_to = "Pathway", values_to = "Score") %>%
  group_by(Pathway) %>%
  mutate(Score_Z = as.numeric(scale(Score))) %>%
  ungroup()

filtered_pathway_matrix_hgs <- filtered_pathway_summary_hgs %>%
  dplyr::select(SampleID, Pathway, Score_Z) %>%
  pivot_wider(names_from = Pathway, values_from = Score_Z) %>%
  column_to_rownames("SampleID") %>%
  as.matrix()
```
#### (hgs 1)
```{r}
#filtered_pathway_summary_hgsc <- filtered_pathway_summary_hgs[hgsc_lines, ]

h1 <- Heatmap(filtered_pathway_matrix_hgs,
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = TRUE,
              cluster_rows = TRUE,
              show_column_dend = TRUE,
              clustering_method_rows = "ward.D2",
              show_row_names = TRUE,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(filtered_pathway_matrix_hgs) * unit(5, "mm"), 
              height = nrow(filtered_pathway_matrix_hgs) * unit(5, "mm"))

pdf(file = "../figs2/progeny_heatmap_RNA_h1_hgs.pdf", width = 15, height = 15)
draw(h1, auto_adjust = FALSE)
dev.off()
```
#### (hgs 2)
```{r}
h2 <- Heatmap(filtered_score_matrix_hgs[hgsc_lines, meta_pathway],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = TRUE,
              cluster_rows = TRUE,
              show_column_dend = TRUE,
              clustering_method_rows = "ward.D2",
              show_row_names = TRUE,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(filtered_score_matrix_hgs[hgsc_lines, meta_pathway]) * unit(5, "mm"), 
              height = nrow(filtered_score_matrix_hgs[hgsc_lines, meta_pathway]) * unit(5, "mm"))

pdf(file = "../figs2/progeny_heatmap_RNA_h2_hgs.pdf", width = 15, height = 15)
draw(h2, auto_adjust = FALSE)
dev.off()

```
#### (hgs 3)

```{r}
h3 <- Heatmap(filtered_score_matrix_hgs[hgsc_lines, functions],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = TRUE,
              cluster_rows = TRUE,
              show_column_dend = TRUE,
              clustering_method_rows = "ward.D2",
              show_row_names = TRUE,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(filtered_score_matrix_hgs[hgsc_lines, functions]) * unit(5, "mm"), 
              height = nrow(filtered_score_matrix_hgs[hgsc_lines, functions]) * unit(5, "mm"))

pdf(file = "../figs2/progeny_heatmap_RNA_h3_hgs.pdf", width = 15, height = 15)
draw(h3, auto_adjust = FALSE)
dev.off()

```
####(hgs 4)
```{r}
h4 <- Heatmap(filtered_score_matrix_hgs[hgsc_lines, metabolism],
              name = "Relative\nactivity",
              col = col_fun,
              heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
              cluster_columns = TRUE,
              cluster_rows = FALSE,
              show_column_dend = TRUE,
              clustering_method_rows = "ward.D2",
              show_row_names = TRUE,
              column_names_rot = 45,
              column_names_gp = gpar(fontsize = 10),
              border = TRUE,
              rect_gp = gpar(col = "white", lwd = 1),
              width = ncol(filtered_score_matrix_hgs[hgsc_lines, metabolism]) * unit(5, "mm"), 
              height = nrow(filtered_score_matrix_hgs[hgsc_lines, metabolism]) * unit(5, "mm"))

pdf(file = "../figs2/progeny_heatmap_RNA_h4_hgs.pdf", width = 15, height = 15)
draw(h4, auto_adjust = FALSE)
dev.off()

```
# ADC targets expression heatmap
```{r}
adc_antigens <- c("FOLR1", "MSLN", "TACSTD2", "ERBB2", "TF", "DPEP3", "SLC34A2")
adc_subset <- intersect(adc_antigens, rownames(tpm_mat_log))
adc_expr <- tpm_mat_log[adc_subset, ]
ordered_samples <- annotation %>% arrange(Subtype) %>% pull(cell_line) 
adc_expr <- adc_expr[, ordered_samples]
adc_scaled <- t(scale(t(adc_expr)))
adc_scaled[adc_scaled > 2] <- 2
adc_scaled[adc_scaled < -2] <- -2

sub_anno <- data.frame(Subtype=annotation$Subtype)
rownames(sub_anno) <- annotation$cell_line
sub_anno <- sub_anno[ordered_samples, , drop = FALSE]

subtypes <- unique(sub_anno$Subtype)
subtype_colors <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
anno_colors <- list(Subtype = subtype_colors)

ht <- Heatmap(
  adc_scaled,
  name = "Z-score",
  bottom_annotation = HeatmapAnnotation(
    Subtype = sub_anno$Subtype,
    col = list(Subtype = subtype_colors),
    show_annotation_name = TRUE
  ),
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 6),
  row_names_gp = gpar(fontsize = 8),
  heatmap_legend_param = list(title = "Expression\nZ-score", title_gp = gpar(fontsize = 10)),
  border = TRUE
)
ht

# 
pdf(file = "../figs2/rna_antigen_heatmap_NRC_CAR.pdf", width = 6, height = 4)
draw(ht, auto_adjust = FALSE)
dev.off()


```
# ADC targets expression tpm + z score dot plot
```{r}
# ADC antigen genes
adc_antigens <- c("FAP", "EGFR", "MSLN", "AXL", "ERBB2", "CA9", "MUC1", "CEACAM6", "CDH17")
adc_subset <- intersect(adc_antigens, rownames(tpm_mat_log))

# Subset TPM matrix
adc_expr <- tpm_mat_log[adc_subset, ]
ordered_samples <- annotation %>% arrange(Subtype) %>% pull(cell_line)
adc_expr <- adc_expr[, ordered_samples]

# Z-score per gene (for color)
adc_scaled <- t(scale(t(adc_expr)))
adc_scaled[adc_scaled > 2] <- 2
adc_scaled[adc_scaled < -2] <- -2

# Make long-format dataframe for plotting
plot_df <- expand.grid(
  Gene = rownames(adc_expr),
  Sample = colnames(adc_expr)
)
plot_df$TPM <- as.vector(adc_expr)
plot_df$Zscore <- as.vector(adc_scaled)

# Add subtype info
plot_df <- plot_df %>%
  left_join(annotation, by = c("Sample" = "cell_line"))

# Factor levels for consistent ordering
plot_df$Gene <- factor(plot_df$Gene, levels = adc_subset)
plot_df$Sample <- factor(plot_df$Sample, levels = ordered_samples)

# Dot plot visualization
dot_plot <- ggplot(plot_df, aes(x = Gene, y = Sample)) +
  geom_point(aes(size = TPM, fill = Zscore), shape = 21, color = "black") +
  scale_size("TPM", range = c(0.5, 6)) +
 scale_fill_gradientn(
  colours = grDevices::colorRampPalette(c("navy","white","firebrick"))(100),
  guide = guide_colorbar(ticks.colour = "black", frame.colour = "black"),
  name = "Expression\nZ-score"
) +
  ylab("") + xlab("") + ggtitle("ADC Target Expression") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1, color = "black"),
        axis.text.y = element_text(size = 7, color = "black"),
        plot.title = element_text(size = 14, hjust = 0.5))

#facet by Subtype

subtype_labs <- c(
  "High-grade serous" = "HGS",
  "Clear cell" = "CC",
  "Endometrioid" = "EC",
  "Mucinous" = "MC",
  "MMMT" = "MMMT",
  "SCCOHT" = "SCCOHT"
)
dot_plot <- dot_plot +
  facet_grid(
    Subtype ~ .,
    scales = "free_y",
    space = "free_y",
    labeller = labeller(Subtype = subtype_labs)
  )
dot_plot
pdf(file = "../figs2/adc_antigen_dotplot_NRC_CAR_RNA.pdf", width = 6, height = 10)
print(dot_plot)
dev.off()
```

# hgs only ov concensus
```{r}

library(consensusOV)
library(org.Hs.eg.db) 

# subset for HGSC samples only 
hgsc_samples <- c(
  "OV1369-R2",
  "OV1946",
  "OV2085",
  "OV2295",
  "OV2295-R2",
  "OV3133-R",
  "OV3331",
  "OV4453",
  "OV4485",
  "OV866-2",
  "OV90",
  "TOV1369",
  "TOV3041G",
  "TOV3133G",
  "TOV3291G"
)
tpm_mat_log <- as.matrix(tpm_mat_log)

# Map gene symbols to Entrez IDs
symbols <- rownames(tpm_mat_log)
entrez <- mapIds(
  org.Hs.eg.db,
  keys    = symbols,
  keytype = "SYMBOL",
  column  = "ENTREZID"
)

# Keep only mapped genes
keep <- !is.na(entrez)
expr_mat   <- tpm_mat_log[keep, , drop = FALSE]
entrez_ids <- as.character(entrez[keep])

#HGSC samples only
expr_mat_hgsc <- expr_mat[, colnames(expr_mat) %in% hgsc_samples, drop = FALSE]

# consensusOV subtyping
subtypes_hgsc <- get.subtypes(
  expression.dataset = expr_mat_hgsc,
  entrez.ids         = entrez_ids,
  method             = "consensusOV"
)

#   table
subtype_assignments_hgsc <- data.frame(
  sample  = colnames(expr_mat_hgsc),
  subtype = subtypes_hgsc$consensusOV.subtypes,
  row.names = NULL
)


write.csv( subtype_assignments_hgsc, file = "/Volumes/Eevee/projects/ovcan_human/output/subtype_assignments_HGSC.csv",row.names = FALSE)

```


# Check for known marker genes for each molecular HGSC subtype
```{r}
marker_genes <- c("CXCL11","CXCL10","CXCR3","HMGA2",
                  "SOX11","MCM2","PCNA","MUC16",
                  "MUC1","SLPI","HOXA9","HOXA10", "HOXA11", "FAP", "ANGPTL2","ANGPTL1")

m <- expr_mat_hgsc[rownames(expr_mat_hgsc) %in% marker_genes, , drop = FALSE]
pdf("/Volumes/Eevee/projects/ovcan_human/figs2/HGSC_CONSOV_markers.pdf", width = 7, height = 9)
pheatmap::pheatmap(m)
dev.off()

```


#final heatmap for consensus marker expression with annotation bar
```{r}
library(pheatmap)

marker_genes <- c("CXCL11","CXCL10","CXCR3","HMGA2",
                  "SOX11","MCM2","PCNA","MUC16",
                  "MUC1","SLPI","HOXA9","HOXA10","HOXA11",
                  "FAP","ANGPTL2","ANGPTL1")

m <- expr_mat_hgsc[rownames(expr_mat_hgsc) %in% marker_genes, , drop = FALSE]

subtype_map <- c(
  "OV1369-R2" = "Differentiated",
  "OV1946"    = "Mesenchymal",
  "OV2085"    = "Immunoreactive",
  "OV2295"    = "Differentiated",
  "OV2295-R2" = "Immunoreactive",
  "OV3133-R"  = "Differentiated",
  "OV3331"    = "Proliferative",
  "OV4453"    = "Proliferative",
  "OV4485"    = "Differentiated",
  "OV866-2"   = "Mesenchymal",
  "OV90"      = "Immunoreactive",
  "TOV1369"   = "Differentiated",
  "TOV3041G"  = "Mesenchymal",
  "TOV3133G"  = "Differentiated",
  "TOV3291G"  = "Mesenchymal"
)

ann_col <- data.frame(
  Subtype = factor(subtype_map[colnames(m)],
                   levels = c("Differentiated","Immunoreactive","Mesenchymal","Proliferative")),
  row.names = colnames(m),
  check.names = FALSE
)

ann_colors <- list(
  Subtype = c(
    Differentiated  = "#1b9e77",
    Immunoreactive  = "#7570b3",
    Mesenchymal     = "#d95f02",
    Proliferative   = "#e7298a"
  )
)

pdf("/Volumes/Eevee/projects/ovcan_human/figs2/HGSC_CONS_OV_marker_heatmap.pdf",
    width = 7, height = 7)

ph <- pheatmap::pheatmap(
  m,                        # raw TPM
  annotation_col = ann_col, # shows at TOP
  annotation_colors = ann_colors,
  annotation_names_col = TRUE,
  silent = TRUE
)
grid::grid.newpage()
grid::grid.draw(ph$gtable)

dev.off()
```

