---
title: "All inclusive final code Protein"
output: html_notebook
---

# Dependencies
```{r}
library(tximport)
library(tidyverse)
library(Hmisc)
library(factoextra)
library(pheatmap)
library(ggrepel)
library(biomaRt)
library(rhdf5)
library(matrixStats)
library(GO.db)
library(Rtsne)     
library(topGO)
library(ComplexHeatmap)
library(readxl)
library(readr)
library(cluster)
library(dplyr)
library(ggplot2)
library(viridisLite)
```

# Load the data
```{r}
raw_data <- read_excel("../data/proteomics/protein_relative_abundance.xlsx")

```
 
# Subset for information we need 
```{r}
file_path <- "../data/proteomics/protein_relative_abundance.xlsx"  
df <- read_excel(file_path, sheet = 1)  

# columns i want
columns_to_keep <- c("OV1369_R2", "OV3133_R", "OV866_2", "VOA10816", 
                     "TOV3392D", "VOA4395", "COV434", "OV2085", "OV4485", 
                     "TOV1369", "VOA295", "VOA12539", "TOV3291G", 
                     "VOA10816.1", "OV2295_R2", "BIN67", "TOV3041G", 
                     "VOA14993", "VOA4841", "TOV1369.1", 
                     "OV4453", "OV3331", "TOV3133G", 
                     "TOV112D", "VOA8762", "VOA5217", 
                     "OV1946", "OV90", "OV2295", "TOV21G", "TOV2414", "VOA5436", 
                     "VOA8771", "TOV3133G.1")

# extract 2 wanted columns
abundance_df <- df[, c("Symbol", columns_to_keep)]

write.csv(abundance_df, "../output/filtered_abundance_table.csv", row.names = FALSE)

```

# Turn CSV -> Matrix
```{r}
abundance_df <- read.csv("../output/filtered_abundance_table.csv")
abundance_df <- abundance_df[!is.na(abundance_df$Symbol), ]
abundance_df$Symbol <- make.unique(as.character(abundance_df$Symbol))
rownames(abundance_df) <- abundance_df$Symbol
abundance_df$Symbol <- NULL


abundance_matrix <- as.matrix(abundance_df)
abundance_matrix <- na.omit(abundance_df)


write.csv(abundance_matrix, "../output/abundance_matrix.csv", row.names = FALSE)

```

# Annotation table 
```{r}
protein_annotation_df_new <- data.frame(
  CellLine = c(
    "OV1369_R2", "OV2085", "TOV3291G", "OV2295_R2", "OV4453", 
    "OV3133_R", "OV4485", "TOV3041G", "OV1946", "TOV3133G", 
    "OV866_2", "TOV1369", "TOV1369.1", "OV2295", "TOV3133G.1", 
    "VOA295", "VOA10816.1", "VOA10816", "VOA12539", 
    "TOV21G", "TOV3392D", "VOA4841", "VOA14993", "VOA4395", "TOV112D", 
    "COV434", "BIN67", "OV3331", "OV90", "TOV2414", 
    "VOA8771", "VOA8762", "VOA5217", "VOA5436"
  ),
  Subtype = c(
    "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", 
    "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", 
    "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", "High-grade serous", "Clear cell", "Clear cell", "Clear cell", "Clear cell", 
    "Clear cell", "Clear cell", "Clear cell", "Clear cell", "Endometrioid", "Endometrioid", 
    "SCCOHT", "SCCOHT", "High-grade serous", "High-grade serous", "Mucinous", 
    "Mucinous", "Mucinous", "MMMT", "MMMT"
  )
)

write.csv(protein_annotation_df_new, "../output/protein_annotation_df.csv", row.names = FALSE)
```



# RNA vs Protein plot

```{r}

protein_matrix <- read.csv("../output/abundance_matrix_best.csv", row.names = 1)
rna_matrix <- read.csv("../output/tpm_matrix_human.csv", row.names = 1)
rna_matrix <- log2(rna_matrix + 1)

abundance_matrix <- protein_matrix

# find the common gene symbols. use intersect tool
common_genes <- intersect(rownames(rna_matrix), rownames(protein_matrix))
# extract the rows corresponding to the gene symbols
rna_common <- rna_matrix[common_genes, ]
protein_common <- protein_matrix[common_genes, ]

# find common cell lines
common_lines <- intersect(colnames(rna_matrix), colnames(protein_matrix))

#extract the columns corresponding to the cell lines
rna_common <- rna_common[, common_lines]
protein_common <- protein_common[, common_lines]

```
## Visualize 
```{r}

line <- "TOV21G"

df <- data.frame(
  RNA = as.numeric(rna_common[, line]),
  Protein = as.numeric(protein_common[, line]))

R <- cor(df$RNA, df$Protein, use = "complete.obs")

p <- ggplot(df, aes(x = RNA, y = Protein)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(
    title = line,
    x = "RNA Expression (log(TPM+1))",
    y = "Relative Protein Abundance \n (log-normalized)"
  ) +
  annotate("text", x = 0, y = max(df$Protein, na.rm = TRUE),
           label = paste0("R= ", round(R, 2)),
           hjust = 0, vjust = 1, size = 7) +
  theme_classic() +
theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.title = element_text(size = 18),
    axis.text  = element_text(size = 18)
  )
ggsave("../figs2/rna_vs_protein_TOV21G.pdf", p, width = 4.5, height = 4.5, dpi = 300)

```


#PCA prep
```{r}
abundance_matrix <- na.omit(abundance_df)

# Explictly set matrix column order to match protein_annotation
abundance_matrix <- abundance_matrix[,protein_annotation_df_new$CellLine]

pca_result_protein <- prcomp(t(abundance_matrix), scale = FALSE, center=TRUE)

# Base PCA visualization
# test <- fviz_pca_ind(pca_result_protein, axes=c(1,2),
#              repel = TRUE,
#              habillage = protein_annotation_df_new$Subtype,
#              max.overlaps=100)
# test



```

# Subtype annotation
```{r}
subtypes <- unique(sub_anno$Subtype)
subtype_colours <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
anno_colours <- list(Subtype = subtype_colours)
```


# PCA
```{r}
# Extract PCA coordinates
plot_dat <- protein_annotation_df_new
plot_dat$PC1 <- pca_result_protein$x[, 1]
plot_dat$PC2 <- pca_result_protein$x[, 2]
plot_dat$PC3 <- pca_result_protein$x[,3]
plot_dat$PC4 <- pca_result_protein$x[,4]
plot_dat$PC5 <- pca_result_protein$x[,5]
plot_dat$PC6 <- pca_result_protein$x[,6]
plot_dat$PC7 <- pca_result_protein$x[,7]
plot_dat$PC8 <- pca_result_protein$x[,8]
plot_dat$PC9 <- pca_result_protein$x[,9]
plot_dat$PC10 <- pca_result_protein$x[,10]


# Plot
pca_plot_protein <- ggplot(plot_dat, aes(x = PC1, y = PC2, label = CellLine)) + 
  geom_point(shape = 21, size = 2, color = "black", aes(fill = Subtype)) +
  geom_text_repel(max.overlaps = 100, size = 1.75) +
  scale_fill_manual(values = subtype_colours) +
  theme_classic() +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)) +
  labs(x = "PC1 (14.2% Variance)", y = "PC2 (11% Variance)")
ggsave("../figs2/pca_plot_protein.pdf", plot = pca_plot_protein, width = 4.5, height = 2.75, dpi = 300)

```

# TSNE
```{r}
ovcan_tsne_protein <- Rtsne::Rtsne(pca_result_protein$x[,1:5], perplexity=5) 

plot_dat <- protein_annotation_df_new
plot_dat$tSNE1 <- ovcan_tsne_protein$Y[,1]
plot_dat$tSNE2 <- ovcan_tsne_protein$Y[,2]

tsne_plot_protein <- ggplot(plot_dat, aes(x=tSNE1, y= tSNE2, label=CellLine)) + 
  geom_point(shape=21, size=2, color="black", aes(fill=Subtype)) +
  geom_text_repel(max.overlaps=100, size=1.75) +
  scale_fill_manual(values = subtype_colours) +
  theme_classic() +
  theme(axis.text=element_text(size=10, color="black"),
        axis.title=element_text(size=12))
ggsave("../figs2/tsne_plot_protein.pdf", plot = tsne_plot_protein, width = 4.5, height = 2.75, dpi = 300)
```



# Silhouette Width PCA
```{r}
dists_protein <- dist(plot_dat[, c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10")])

# Cluster labels from subtypes
cluster_labels_protein <- as.numeric(factor(plot_dat$Subtype))

# Compute silhouette values
sil_protein <- silhouette(cluster_labels_protein, dists_protein)
sil_protein_df <- as.data.frame(sil_protein)
sil_protein_df$Subtype <- plot_dat$Subtype
sil_summary_protein <- sil_protein_df %>%
  group_by(Subtype) %>%
  summarise(mean_sil = mean(sil_width)) %>%
  arrange(mean_sil)

sil_lollipop_protein <- ggplot(sil_summary_protein, aes(x = mean_sil, y = reorder(Subtype, mean_sil))) +
  geom_bar(stat = "identity", width = 0.01, color = "black") +
  geom_point(fill = "#4A154B", shape = 21, size = 4, color = "black") +
  geom_vline(xintercept = 0, linetype = 2) +
  ggtitle("Silhouette Width by Subtype (Protein PCA)") +
  xlab("Mean Silhouette Width") + ylab("") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 26),
    plot.title = element_text(size = 26, face = "bold")
  )
ggsave("../figs2/silhouette_lollipop_protein_1-10.pdf", plot = sil_lollipop_protein, width = 7.5, height = 6)

```

# Signature heatmap 
```{r}
rownames(abundance_df) <- abundance_df$Symbol

common_cell_lines <- intersect(colnames(tpm_mat_log), colnames(abundance_df))
signature_genes <- unique(df_sig_hm$Gene)  
shared_genes <- intersect(signature_genes, rownames(abundance_df))
protein_sig_mat <- abundance_df[shared_genes, common_cell_lines]

# col anno
sample_anno <- read.csv("../output/annotation.csv") %>%
  filter(cell_line %in% colnames(protein_sig_mat)) %>%
  arrange(Subtype)  # Reorder by subtype

# Reorder columns of the matrix
protein_sig_mat <- protein_sig_mat[, sample_anno$cell_line]
protein_sig_mat <- na.omit(protein_sig_mat)

protein_scaled <- t(scale(t(protein_sig_mat)))
protein_scaled[protein_scaled > 3] <- 3
protein_scaled[protein_scaled < -3] <- -3

subtype_vector_gene <- df_sig_hm$Subtype
names(subtype_vector_gene) <- df_sig_hm$Gene # gene to subtype mapping for annotation

# keeps only the genes that are plotted in the heatmap and puts them in the same order, so the annotation colours line up with the rows
subtype_vector_gene <- subtype_vector_gene[rownames(protein_scaled)]

row_anno <- rowAnnotation(
  Subtype = subtype_vector_gene,
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE)

bottom_anno <- HeatmapAnnotation(
  Subtype = subtype_vector_gene,
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE)

# Visualization
sigs_protein_heatmap <- Heatmap(
  protein_scaled,
  name = "Expression\nZ-score",
  show_row_names = TRUE,
  show_column_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 6),
  heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
  border = TRUE,
  left_annotation = row_anno,
  bottom_annotation = bottom_anno)

pdf(file = "../figs2/sigs_heatmap_protein_top100.pdf", width = 8, height = 8)
draw(sigs_protein_heatmap)
dev.off()
```



# ADC target heatmap
```{r}
adc_antigens <- c("FOLR1", "MSLN", "TACSTD2", "ERBB2", "TF", "DPEP3", "SLC34A2")
adc_subset <- intersect(adc_antigens, rownames(abundance_matrix))
adc_expr <- abundance_matrix[adc_subset, ]
ordered_samples <- protein_annotation_df_new %>% arrange(Subtype) %>% pull(CellLine) 
adc_expr <- adc_expr[, ordered_samples]
adc_matrix <- as.matrix(adc_expr)

adc_scaled <- t(base::scale(t(adc_matrix)))

adc_scaled[adc_scaled > 2] <- 2
adc_scaled[adc_scaled < -2] <- -2


adc_scaled_t <- t(adc_scaled)  # transpose the matrix

sub_anno <- data.frame(Subtype=protein_annotation_df_new$Subtype)
rownames(sub_anno) <- protein_annotation_df_new$CellLine
sub_anno <- sub_anno[ordered_samples, , drop = FALSE]

subtypes <- unique(sub_anno$Subtype)
subtype_colors <- setNames(brewer.pal(length(subtypes), "Set2"), subtypes)
anno_colors <- list(Subtype = subtype_colors)

ht <- Heatmap(
  adc_scaled_t,
  name = "Z-score",
  right_annotation = HeatmapAnnotation(
    Subtype = sub_anno$Subtype,
    col = list(Subtype = subtype_colors),
    show_annotation_name = TRUE,
    which = "row"
  ),
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  column_names_gp = gpar(fontsize = 11.5),
  row_names_gp = gpar(fontsize = 8),
  heatmap_legend_param = list(title = "Expression\nZ-score", title_gp = gpar(fontsize = 10)),
  border = TRUE
)
ht


pdf(file = "../figs2/protein_antigen_heatmap_paper.pdf", width = 4.5, height = 10)
draw(ht, auto_adjust = FALSE)
dev.off()

```

# ADC target dot plot z score + abundance 
```{r}

adc_antigens <- c("FAP", "EGFR", "MSLN", "AXL", "ERBB2", "CA9", "MUC1", "CEACAM6", "CDH17")
adc_subset <- intersect(adc_antigens, rownames(abundance_matrix))

ordered_samples <- protein_annotation_df_new %>%
  arrange(Subtype) %>%
  pull(CellLine)

# IMPORTANT: make it a numeric matrix
adc_expr <- abundance_matrix[adc_subset, ordered_samples, drop = FALSE]
adc_expr <- as.matrix(adc_expr)
storage.mode(adc_expr) <- "numeric"

# Z-score per gene
adc_scaled <- t(scale(t(adc_expr)))
adc_scaled[adc_scaled > 2] <- 2
adc_scaled[adc_scaled < -2] <- -2

# Long df
plot_df <- expand.grid(
  Gene = rownames(adc_expr),
  Sample = colnames(adc_expr),
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

plot_df$Abundance <- as.vector(adc_expr)
plot_df$Zscore <- as.vector(adc_scaled)

plot_df <- plot_df %>%
 dplyr::left_join(
    protein_annotation_df_new %>% dplyr::select(CellLine, Subtype),
    by = c("Sample" = "CellLine")
  )

plot_df$Gene <- factor(plot_df$Gene, levels = adc_subset)
plot_df$Sample <- factor(plot_df$Sample, levels = ordered_samples)

subtype_labs <- c(
  "High-grade serous" = "HGS",
  "Clear cell" = "CC",
  "Endometrioid" = "EC",
  "Mucinous" = "MC",
  "Low-Grade Serous" = "LGS"
)

protein_dotplot <- ggplot(plot_df, aes(x = Gene, y = Sample)) +
  geom_point(aes(size = Abundance, fill = Zscore), shape = 21, color = "black") +
  scale_size("Protein abundance", range = c(0.5, 6)) +
  scale_fill_gradientn(
  colours = grDevices::colorRampPalette(c("navy","white","firebrick"))(100),
  guide = guide_colorbar(ticks.colour = "black", frame.colour = "black"),
  name = "Expression\nZ-score"
) +
  facet_grid(Subtype ~ ., scales = "free_y", space = "free_y",
             labeller = labeller(Subtype = subtype_labs)) +
  labs(title = "ADC Target protein abundance", x = "", y = "") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1, color = "black"),
        axis.text.y = element_text(size = 7, color = "black"),
        plot.title = element_text(size = 14, hjust = 0.5))

protein_dotplot
pdf(file = "../figs2/adc_antigen_dotplot.pdf", width = 6, height = 10)
print(protein_dotplot)
dev.off()

```

# t test to identify molecular signatures 
```{r}

gene_abundance <- as.numeric(abundance_matrix[gene, ])

cell_lines <- colnames(abundance_matrix)


target_subtype <- "SCCOHT"

abundance_clean <- na.omit(abundance_matrix)
results <- data.frame( gene = character(),
                       p_value = numeric(),
                       log2fc = numeric(),
                       stringsAsFactors = F)

for (gene in rownames(abundance_clean)) {
  gene_abundance <- as.numeric(abundance_clean[gene, ])
  
df <- data.frame(
    gene = gene,
    abundance = gene_abundance,
    CellLine = cell_lines)
  
df <- merge(df, annotation_table, by = "CellLine")#, all.x = TRUE)

group1 <- df[df$Subtype == target_subtype, "abundance"]
group2 <- df[df$Subtype != target_subtype, "abundance"]

#if (sum(!is.na(group1)) < 1 || sum(!is.na(group2)) < 1) next

test <- t.test(group1, group2, var.equal = TRUE)

log2fc <- mean(group1, na.rm = TRUE) - mean(group2, na.rm = TRUE)

  results <- rbind(results, data.frame(
    Gene = gene,
    p_value = test$p.value,
    log2FoldChange = log2fc))
}

results$padj <- p.adjust(results$p_value, method = "BH") #controls false discovery rate across all genes tested
write.csv(results, paste0("ttest", gsub(" ", "_", target_subtype), "_vs_OTHERS.csv"), row.names = FALSE)

```

# Signature for Clear Cell with pajd/log2fc using t-test
```{r}
cc_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestClear_cell_vs_OTHERS.csv", row.names = 1)
cc_signature_df_protein <- subset(cc_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
cc_signature_df_protein$Gene <- rownames(cc_signature_df_protein)
cc_signature_df_protein <- cc_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(cc_signature_df_protein,
          "../output/Protein signatures/ttest_sig/ClearCell_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```

# Signature for High-Grade with pajd/log2fc using t-test
```{r}
hgs_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestHigh-grade_serous_vs_OTHERS.csv", row.names = 1)
hgs_signature_df_protein <- subset(hgs_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
hgs_signature_df_protein$Gene <- rownames(hgs_signature_df_protein)
hgs_signature_df_protein <- hgs_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(hgs_signature_df_protein,
          "../output/Protein signatures/ttest_sig/HGS_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```

# Signature for Mucinous with pajd/log2fc using t-test
```{r}
mc_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestMucinous_vs_OTHERS.csv", row.names = 1)
mc_signature_df_protein <- subset(mc_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
mc_signature_df_protein$Gene <- rownames(mc_signature_df_protein)
mc_signature_df_protein <- mc_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(mc_signature_df_protein,
          "../output/Protein signatures/ttest_sig/Mucinous_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```

# Signature for Endometrioid with pajd/log2fc using t-test
```{r}
ec_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestEndometrioid_vs_OTHERS.csv", row.names = 1)
ec_signature_df_protein <- subset(ec_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
ec_signature_df_protein$Gene <- rownames(ec_signature_df_protein)
ec_signature_df_protein <- ec_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(ec_signature_df_protein,
          "../output/Protein signatures/ttest_sig/Endometrioid_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```


# Signature for SCCOHT with pajd/log2fc using t-test
```{r}
SCCOHT_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestSCCOHT_vs_OTHERS.csv", row.names = 1)
SCCOHT_signature_df_protein <- subset(SCCOHT_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
SCCOHT_signature_df_protein$Gene <- rownames(SCCOHT_signature_df_protein)
SCCOHT_signature_df_protein <- SCCOHT_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(SCCOHT_signature_df_protein,
          "../output/Protein signatures/ttest_sig/SCCOHT_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```

# Signature for MMMT with pajd/log2fc using t-test
```{r}
MMMT_protein <- read.csv("/Volumes/Eevee/projects/ovcan_human/output/ttest/ttestMMMT_vs_OTHERS.csv", row.names = 1)
MMMT_signature_df_protein <- subset(MMMT_protein, padj < 0.05 & log2FoldChange > 1) %>%
  arrange(padj)
MMMT_signature_df_protein$Gene <- rownames(MMMT_signature_df_protein)
MMMT_signature_df_protein <- MMMT_signature_df_protein[, c("Gene", "padj", "log2FoldChange")]
write.csv(MMMT_signature_df_protein,
          "../output/Protein signatures/ttest_sig/MMMT_signature_protein.csv",
          row.names = FALSE,
          quote = FALSE)

```

# find common signature between protein and rna (HGS)
```{r}
hgs_sig_p <- read_csv("../output/Protein signatures/ttest_sig/HGS_signature_protein.csv")
hgs_sig_r <- read_csv("../output/HGS_signature_detailed.csv")

common_sig <- intersect(hgs_sig_p$Gene, hgs_sig_r$Gene)

hgs_merged_sig <- merge(hgs_sig_p[hgs_sig_p$Gene %in% common_sig, ], hgs_sig_r[hgs_sig_r$Gene %in% common_sig, ], by = "Gene")

write.csv(hgs_merged_sig, "../output/hgs_merged_sig.csv", row.names = FALSE)

```


# find common signature between protein and rna (Mucinous)
```{r}
mc_sig_p <- read_csv("../output/Protein signatures/ttest_sig/Mucinous_signature_protein.csv")
mc_sig_r <- read_csv("../output/Mucinous_signature_detailed.csv")

common_sig_mc <- intersect(mc_sig_p$Gene, mc_sig_r$Gene)

mc_merged_sig <- merge(mc_sig_p[mc_sig_p$Gene %in% common_sig_mc, ], mc_sig_r[mc_sig_r$Gene %in% common_sig_mc, ], by = "Gene")

write.csv(mc_merged_sig, "../output/mc_merged_sig.csv", row.names = FALSE)

```
# find common signature between protein and rna (Clear cell)
```{r}
cc_sig_p <- read_csv("../output/Protein signatures/ttest_sig/ClearCell_signature_protein.csv")
cc_sig_r <- read_csv("../output/ClearCell_signature_detailed.csv")

common_sig_cc <- intersect(cc_sig_p$Gene, cc_sig_r$Gene)

cc_merged_sig <- merge(cc_sig_p[cc_sig_p$Gene %in% common_sig_cc, ], cc_sig_r[cc_sig_r$Gene %in% common_sig_cc, ], by = "Gene")

write.csv(cc_merged_sig, "../output/cc_merged_sig.csv", row.names = FALSE)

```

# find common signature between protein and rna (Endometrioid)
```{r}
ec_sig_p <- read_csv("../output/Protein signatures/ttest_sig/Endometrioid_signature_protein.csv")
ec_sig_r <- read_csv("../output/ec_signature_detailed.csv")

common_sig_ec <- intersect(ec_sig_p$Gene, ec_sig_r$Gene)

ec_merged_sig <- merge(ec_sig_p[ec_sig_p$Gene %in% common_sig_ec, ], ec_sig_r[ec_sig_r$Gene %in% common_sig_ec, ], by = "Gene")

write.csv(ec_merged_sig, "../output/ec_merged_sig.csv", row.names = FALSE)

```

# find common signature between protein and rna (SCCOHT)
```{r}
sccoht_sig_p <- read_csv("../output/Protein signatures/ttest_sig/SCCOHT_signature_protein.csv")
sccoht_sig_r <- read_csv("../output/sccoht_signature_detailed.csv")

common_sig_sccoht <- intersect(sccoht_sig_p$Gene, sccoht_sig_r$Gene)

sccoht_merged_sig <- merge(sccoht_sig_p[sccoht_sig_p$Gene %in% common_sig_sccoht, ], sccoht_sig_r[sccoht_sig_r$Gene %in% common_sig_sccoht, ], by = "Gene")

write.csv(sccoht_merged_sig, "../output/sccoht_merged_sig.csv", row.names = FALSE)

```

# find common signature between protein and rna (MMMT)
```{r}
MMMT_sig_p <- read_csv("../output/Protein signatures/ttest_sig/MMMT_signature_protein.csv")
MMMT_sig_r <- read_csv("../output/MMMT_signature_detailed.csv")

common_sig_MMMT <- intersect(MMMT_sig_p$Gene, MMMT_sig_r$Gene)

MMMT_merged_sig <- merge(MMMT_sig_p[MMMT_sig_p$Gene %in% common_sig_MMMT, ], MMMT_sig_r[MMMT_sig_r$Gene %in% common_sig_MMMT, ], by = "Gene")

write.csv(MMMT_merged_sig, "../output/MMMT_merged_sig.csv", row.names = FALSE)

```


# final merged signature data frame 
```{r}
# call out each subtype and its signature files : 


# cc
cc_sig_p <- read_csv("../output/Protein signatures/ttest_sig/ClearCell_signature_protein.csv")
cc_sig_r <- read_csv("../output/ClearCell_signature_detailed.csv")

common_sig_cc <- intersect(cc_sig_p$Gene, cc_sig_r$Gene)

cc_merged_sig <- merge(
  cc_sig_p[cc_sig_p$Gene %in% common_sig_cc, ],
  cc_sig_r[cc_sig_r$Gene %in% common_sig_cc, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
cc_merged_sig$Subtype <- "Clear cell"

# ec 
ec_sig_p <- read_csv("../output/Protein signatures/ttest_sig/Endometrioid_signature_protein.csv")
ec_sig_r <- read_csv("../output/ec_signature_detailed.csv")

common_sig_ec <- intersect(ec_sig_p$Gene, ec_sig_r$Gene)

ec_merged_sig <- merge(
  ec_sig_p[ec_sig_p$Gene %in% common_sig_ec, ],
  ec_sig_r[ec_sig_r$Gene %in% common_sig_ec, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
ec_merged_sig$Subtype <- "Endometrioid"

#hgsc
hgs_sig_p <- read_csv("../output/Protein signatures/ttest_sig/HGS_signature_protein.csv")
hgs_sig_r <- read_csv("../output/HGS_signature_detailed.csv")

common_sig <- intersect(hgs_sig_p$Gene, hgs_sig_r$Gene)

hgs_merged_sig <- merge(
  hgs_sig_p[hgs_sig_p$Gene %in% common_sig, ],
  hgs_sig_r[hgs_sig_r$Gene %in% common_sig, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
hgs_merged_sig$Subtype <- "High-grade serous"

# mmmt
MMMT_sig_p <- read_csv("../output/Protein signatures/ttest_sig/MMMT_signature_protein.csv")
MMMT_sig_r <- read_csv("../output/MMMT_signature_detailed.csv")

common_sig_MMMT <- intersect(MMMT_sig_p$Gene, MMMT_sig_r$Gene)

MMMT_merged_sig <- merge(
  MMMT_sig_p[MMMT_sig_p$Gene %in% common_sig_MMMT, ],
  MMMT_sig_r[MMMT_sig_r$Gene %in% common_sig_MMMT, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
MMMT_merged_sig$Subtype <- "MMMT"

# mc
mc_sig_p <- read_csv("../output/Protein signatures/ttest_sig/Mucinous_signature_protein.csv")
mc_sig_r <- read_csv("../output/Mucinous_signature_detailed.csv")

common_sig_mc <- intersect(mc_sig_p$Gene, mc_sig_r$Gene)

mc_merged_sig <- merge(
  mc_sig_p[mc_sig_p$Gene %in% common_sig_mc, ],
  mc_sig_r[mc_sig_r$Gene %in% common_sig_mc, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
mc_merged_sig$Subtype <- "Mucinous"


# sccoht
sccoht_sig_p <- read_csv("../output/Protein signatures/ttest_sig/SCCOHT_signature_protein.csv")
sccoht_sig_r <- read_csv("../output/sccoht_signature_detailed.csv")

common_sig_sccoht <- intersect(sccoht_sig_p$Gene, sccoht_sig_r$Gene)

sccoht_merged_sig <- merge(
  sccoht_sig_p[sccoht_sig_p$Gene %in% common_sig_sccoht, ],
  sccoht_sig_r[sccoht_sig_r$Gene %in% common_sig_sccoht, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
sccoht_merged_sig$Subtype <- "SCCOHT"


#merge all subtypes
all_merged <- rbind(
  cc_merged_sig,
  ec_merged_sig,
  hgs_merged_sig,
  MMMT_merged_sig,
  mc_merged_sig,
  sccoht_merged_sig
  )

all_merged <- all_merged[, c("Subtype", setdiff(names(all_merged), "Subtype"))]
write.csv(all_merged, "../output/protein_rna_signatures.csv", row.names = FALSE)

```


# Final Molecular signature heatmap using RNA 
 
```{r}

annotation <- read_csv("../output/annotation.csv")
sig_merged <- read_csv("../output/protein_rna_signatures.csv")
sig_genes <- unique(sig_merged$Gene)

tpm_w_sig <- tpm_mat_log[sig_genes, , drop =FALSE]

tpm_w_sig <- tpm_w_sig[rowVars(as.matrix(tpm_w_sig)) != 0, ]
tpm_w_sig <- t(scale(t(tpm_w_sig)))  # Z-score per gene

tpm_w_sig[tpm_w_sig > 2] <- 2
tpm_w_sig[tpm_w_sig < -2] <- -2

sample_anno <- read.csv("../output/annotation.csv") %>%
  filter(cell_line %in% colnames(tpm_w_sig)) %>%
  arrange(Subtype)

tpm_w_sig <- tpm_w_sig[, sample_anno$cell_line]

# Build row subtype vector from sig_merged
subtype_vector_row <- sig_merged$Subtype
names(subtype_vector_row) <- sig_merged$Gene

# subtype annotation
unique_subtypes <- sort(unique(subtype_vector_row))
subtype_colours <- setNames( RColorBrewer::brewer.pal(length(subtype_levels), "Set2"), subtype_levels)

# row annotation
row_anno <- rowAnnotation(
  Subtype = subtype_vector_row[rownames(tpm_w_sig)],
  col = list(Subtype = subtype_colours),
  show_annotation_name = TRUE,
  show_legend = F)

# column annotation
col_anno <- HeatmapAnnotation(
  Subtype = sample_anno$Subtype,
  col = list(Subtype = subtype_colours),
  show_annotation_name = F,
  show_legend = F)

col_fun <- circlize::colorRamp2(
  seq(-2, 2, length.out = 50),
  colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))(50))

# visualize 
rna_merged_sig_heatmap <- Heatmap(
  tpm_w_sig,
  name = "Expression\nZ-score",
  col = col_fun,
  bottom_annotation = col_anno,
  left_annotation = row_anno,
  show_row_names = F,
  show_column_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 12),
  row_names_gp = gpar(fontsize = 9),
  heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
  width = ncol(tpm_w_sig)*unit(5, "mm"), 
  height = nrow(tpm_w_sig)*unit(5, "mm"),
  border = TRUE)

pdf(file = "../figs2/rna_merged_sig_heatmap.pdf", width = 10, height = 8)
draw(rna_merged_sig_heatmap)
dev.off()


```

# Final Molecular signature heatmap using PROTEIN 
```{r}

annotation <- read_csv("../output/annotation.csv")
sig_merged <- read_csv("../output/protein_rna_signatures.csv")
sig_genes <- unique(sig_merged$Gene)

abundance_w_sig <- abundance_matrix[sig_genes, , drop =FALSE]

abundance_w_sig <- abundance_w_sig[rowVars(as.matrix(abundance_w_sig)) != 0, ]
abundance_w_sig <- t(scale(t(abundance_w_sig)))  # Z-score per gene

# z score
abundance_w_sig[abundance_w_sig > 2] <- 2
abundance_w_sig[abundance_w_sig < -2] <- -2

sample_anno <- read.csv("../output/protein_annotation_df.csv") %>%
  filter(CellLine %in% colnames(abundance_w_sig)) %>%
  arrange(Subtype)

abundance_w_sig <- abundance_w_sig[, sample_anno$CellLine]

subtype_vector_row <- sig_merged$Subtype
names(subtype_vector_row) <- sig_merged$Gene

unique_subtypes <- sort(unique(subtype_vector_row))
subtype_colors <- setNames( RColorBrewer::brewer.pal(length(unique_subtypes), "Set2"),unique_subtypes)

row_anno <- rowAnnotation(
  Subtype = subtype_vector_row[rownames(abundance_w_sig)],
  col = list(Subtype = subtype_colors),
  show_annotation_name = TRUE)

col_anno <- HeatmapAnnotation(
  Subtype = sample_anno$Subtype,
  col = list(Subtype = subtype_colors),
  show_annotation_name = F,
  show_legend = F)

col_fun <- circlize::colorRamp2(
  seq(-2, 2, length.out = 50),
  colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))(50))

protein_merged_sig_heatmap <- Heatmap(
  abundance_w_sig,
  name = "Expression\nZ-score",
  col = col_fun,
  bottom_annotation = col_anno,
  left_annotation = row_anno,
  show_row_names = TRUE,
  show_column_names = TRUE,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  column_names_gp = gpar(fontsize = 12),
  row_names_gp = gpar(fontsize = 11),
  heatmap_legend_param = list(title_gp = gpar(fontsize = 10)),
  width = ncol(abundance_w_sig)*unit(5, "mm"), 
  height = nrow(abundance_w_sig)*unit(5, "mm"),
  border = TRUE
)

pdf(file = "../figs2/protein_merged_sig_heatmap1_paper.pdf", width = 10, height = 8)
draw(protein_merged_sig_heatmap)
dev.off()

```
# RNA and Protein all genes in one csv file
```{r}

# cc
cc_p <- read_csv("../output/ttest/ttestClear_cell_vs_OTHERS.csv")
cc_r <- read_csv("../output/differential_expression_Clear cell_vs_Other.csv")
cc_r <- na.omit(cc_r)
names(cc_r)[1] <- "Gene"


common_cc <- intersect(cc_p$Gene, cc_r$Gene)

cc_merged <- merge(
  cc_p[cc_p$Gene %in% common_cc, ],
  cc_r[cc_r$Gene %in% common_cc, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
cc_merged$Subtype <- "Clear cell"

# ec 
ec_p <- read_csv("../output/ttest/ttestEndometrioid_vs_OTHERS.csv")
ec_r <- read_csv("../output/differential_expression_Endometrioid_vs_Other.csv")
names(ec_r)[1] <- "Gene"
common_ec <- intersect(ec_p$Gene, ec_r$Gene)

ec_merged <- merge(
  ec_p[ec_p$Gene %in% common_ec, ],
  ec_r[ec_r$Gene %in% common_ec, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
ec_merged$Subtype <- "Endometrioid"

#hgs
hgs_p <- read_csv("../output/ttest/ttestHigh-grade_serous_vs_OTHERS.csv")
hgs_r <- read_csv("../output/differential_expression_HGS_vs_Other.csv")
names(hgs_r)[1] <- "Gene"
common_hgs <- intersect(hgs_p$Gene, hgs_r$Gene)

hgs_merged <- merge(
  hgs_p[hgs_p$Gene %in% common_hgs, ],
  hgs_r[hgs_r$Gene %in% common_hgs, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
hgs_merged$Subtype <- "High-grade serous"

# mmmt
MMMT_p <- read_csv("../output/ttest/ttestMMMT_vs_OTHERS.csv")
MMMT_r <- read_csv("../output/differential_expression_MMMT_vs_Other.csv")
names(MMMT_r)[1] <- "Gene"
common_MMMT <- intersect(MMMT_p$Gene, MMMT_r$Gene)

MMMT_merged <- merge(
  MMMT_p[MMMT_p$Gene %in% common_MMMT, ],
  MMMT_r[MMMT_r$Gene %in% common_MMMT, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
MMMT_merged$Subtype <- "MMMT"

# mc
mc_p <- read_csv("../output/ttest/ttestMucinous_vs_OTHERS.csv")
mc_r <- read_csv("../output/differential_expression_Mucinous_vs_Other.csv")
names(mc_r)[1] <- "Gene"
common_mc <- intersect(mc_p$Gene, mc_r$Gene)

mc_merged <- merge(
  mc_p[mc_p$Gene %in% common_mc, ],
  mc_r[mc_r$Gene %in% common_mc, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
mc_merged$Subtype <- "Mucinous"


# sccoht
sccoht_p <- read_csv("../output/ttest/ttestSCCOHT_vs_OTHERS.csv")
sccoht_r <- read_csv("../output/differential_expression_SCCOHT_vs_Other.csv")
names(sccoht_r)[1] <- "Gene"
common_sccoht <- intersect(sccoht_p$Gene, sccoht_r$Gene)

sccoht_merged <- merge(
  sccoht_p[sccoht_p$Gene %in% common_sccoht, ],
  sccoht_r[sccoht_r$Gene %in% common_sccoht, ],
  by = "Gene",
  suffixes = c("_Protein", "_RNA"))
sccoht_merged$Subtype <- "SCCOHT"


#merge
all_merged <- rbind(
  cc_merged,
  ec_merged,
  hgs_merged,
  MMMT_merged,
  mc_merged,
  sccoht_merged
  )

all_merged <- all_merged[, c("Subtype", setdiff(names(all_merged), "Subtype"))]
write.csv(all_merged, "../output/protein_rna_genes.csv", row.names = FALSE)

```
# LOG plots
```{r}
df_clearcell <- subset(all_merged, Subtype == "Clear cell")
all_merged <- "../output/protein_rna_genes.csv"
plot(df_clearcell$log2FoldChange_RNA, df_clearcell$log2FoldChange_Protein,
     xlab = "log2FoldChange_RNA",
     ylab = "log2FoldChange_Protein",
     main = "Clear cell log2FoldChange",
     pch = 19, col = "blue")
```
 
```{r}
df_hgs <- subset(all_merged, Subtype == "High-grade serous")

ggplot(subset(df_hgs, Subtype == "High-grade serous"), aes(x = log2FoldChange_RNA, y = log2FoldChange_Protein)) +
  geom_point(color = "blue", size = 0.8) +  # smaller points
  labs(title = "HGS Log2FC") +
  theme_minimal()

```
```{r}
df_cc <- subset(all_merged, Subtype == "Clear cell")

ggplot(subset(df_cc, Subtype == "Clear cell"), aes(x = log2FoldChange_RNA, y = log2FoldChange_Protein)) +
  geom_point(color = "blue", size = 0.8) +  # smaller points
  labs(title = "CC Log2FC") +
  theme_minimal()
```
```{r}

ggplot(subset(all_merged, Subtype == "MMMT"), aes(x = log2FoldChange_RNA, y = log2FoldChange_Protein)) +
  geom_point(color = "blue", size = 0.8) +  # smaller points
  labs(title = "MMMT Log2FC") +
  theme_minimal()
```

